---
title: "Untitled"
output: github_document
author: "Matthew Gust"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Libraries
```{r}
library(tidyverse)
library(pzfx)
library(scales)
library(writexl)
library(purrr)
```

# Plot Info
```{r}
cols <- c("PBS" = "#818181",
          "oHSV" = "#0C5CA1",
          "Trab." = "#C94B27",
          "oHSV + Trab." = "#8A6BAF",
          "oHSV + anti-NK" = "#0F8F5F",
          "oHSV + Trab. + anti-NK" = "#FBAF3F",
          "oHSV + Trab. + control serum" = "#0F8F5F",
          "oHSV + Trab. + anti-NK" = "#FBAF3F",
          "oHSV + Trab. + anti-CD4" = "#0F8F5F",
          "oHSV + Trab. + anti-CD8" = "#0C5CA1",
          "oHSV + Trab. + anti-CD4 + anti-CD8" = "#FBAF3F")


trt_grp_order <-
    c("PBS",
      "oHSV",
      "oHSV + anti-NK",
      "Trab.",
      "oHSV + Trab.",
      "oHSV + Trab. + control serum",
      "oHSV + Trab. + anti-NK",
      "oHSV + Trab. + anti-CD4",
      "oHSV + Trab. + anti-CD8",
      "oHSV + Trab. + anti-CD4 + anti-CD8")
```

# Helper Functions
```{r}
#Create summarized dataframe for plotting
summarized_df <- function(df) {
    #Create calc_resp if that column is not present
    if(!"calc_resp" %in% colnames(df)) {
        df$calc_resp <- rep(NA, nrow(df))
    }
    
    wtrfall_plt <- group_by(df, Treatment, Mouse_num) %>%
        summarize(min = min(relative_tumor_vol),
                  max = max(relative_tumor_vol),
                  Treatment = unique(Treatment),
                  calc_resp = unique(calc_resp),
                  Experiment = unique(Experiment),
                  rtv_at_max_tp = unique(rtv_at_max_tp))

    #Add best overall response columns
    for (i in seq_len(nrow(wtrfall_plt))) {
        if (wtrfall_plt$calc_resp[i] == "PD") {
            wtrfall_plt$BOR1[i] <- wtrfall_plt$max[i]
            wtrfall_plt$BOR2[i] <- wtrfall_plt$max[i]
        } else if (wtrfall_plt$calc_resp[i] == "PR") {
            wtrfall_plt$BOR1[i] <- max(wtrfall_plt$min[i], 10^-1)
            wtrfall_plt$BOR2[i] <- wtrfall_plt$min[i]
        } else if (wtrfall_plt$calc_resp[i] == "SD") {
            wtrfall_plt$BOR1[i] <- wtrfall_plt$rtv_at_max_tp[i]
            wtrfall_plt$BOR2[i] <- wtrfall_plt$max[i]
        } else {
            wtrfall_plt$BOR1[i] <- 10^-1 # Shortened the CR interval
            wtrfall_plt$BOR2[i] <- 10^-1 # Maybe we can add a horizontal line with a label "CR"? # nolint
        }
    }
    return(wtrfall_plt)
}

#Create waterfall plot from summarized dataframe
make_wtrfall_plot <- function(df) {
    df$Treatment <- factor(df$Treatment, levels = trt_grp_order)
    #Create position column to order bars the right way
    df <- arrange(df, desc(BOR1)) %>%
        arrange(Treatment)

    df$x_position <- seq_len(nrow(df))

    # df$x_position <- nrow(df) + 1 - rank(df$BOR1, ties.method = "random")


    #Variable to hold experiment
    experiment <- unique(df$Experiment)
    #Create plot
    plot <- ggplot(df, aes(x = x_position, y = BOR1)) +
        geom_bar(stat = "identity", aes(fill = Treatment), width = 0.7) +
        scale_fill_manual(values = cols) +
        geom_hline(yintercept = 1) +
        geom_hline(yintercept = 0.5, linetype = "dotted") +
        geom_hline(yintercept = 1.25, linetype = "dotted") +
        geom_hline(yintercept = 0.1) + # Added a line that we will relabel as "CR" # nolint
        labs(x = "",
             y = "Best Overall Response", # More accessible axis title
             title = paste(experiment, "28-day Treatment Response")) + # Will need to make this variable-driven for different figure # nolint
        theme(panel.grid.major = element_blank(),
              panel.grid.minor = element_blank(),
              panel.background = element_blank(),
              axis.text.x = element_blank(),
              axis.ticks.x = element_blank(),
              axis.line.x = element_blank(),
              axis.line.y = element_line(color = "black")) +
              scale_y_log10(labels = c("CR", "-50%", "0%", "25%"),
                            breaks = c(0.1, 0.50, 1, 1.25))
}

#Calculate response category
calc_resp <- function(df) {
    df <- df %>%
        group_by(Treatment, Mouse_num) %>%
        mutate(max_timepoint = max(timepoint)) %>%
        filter(timepoint == max_timepoint) %>%
        select(Treatment,
               Mouse_num,
               max_timepoint,
               rtv_at_max_tp = relative_tumor_vol) %>%
        right_join(df, by = c("Treatment", "Mouse_num"))

    group_by(df, Treatment, Mouse_num) %>%
        mutate(calc_resp = if(any(relative_tumor_vol == 0)) {
            "CR"
        } else if(min(relative_tumor_vol) < 0.5) {
            "PR"
        } else if(unique(rtv_at_max_tp) < 1.25) { # nolint
            "SD"
        } else {
            "PD"
        }
    )
}
```


# Analysis

## Reading in all Files
These files were initially prism files that I converted to csv's in order to easily clean and reformat in excel.
```{r}
#Create variable to store different experiment names
experiments <- list.files("Input")
#List to hold each experiment's combined dataframe
dfs_list <- list()
#Loop through experiments to download files
for(i in experiments) {
    #Var to hold sheets
    sheets <- paste("Input", i, list.files(paste0("Input/", i)), sep = "/")

    sheet_df <- data.frame()
    #Loop through all sheets in each experiment
    for (k in sheets) {
        tmp <- read.csv(k, fileEncoding="UTF-8-BOM") # Added the encoding to make it work cross-platform #nolint
        #Rename columns to ID mice (arbitrary labels)
        names(tmp)[!names(tmp) %in% c("timepoint", "Treatment")] <-
            paste("Mouse", seq_len(ncol(tmp) - 2), sep = "_")
        #Add Column to identify which experiment it is
        tmp$Experiment <- i
        #Pivot longer so all Mouse_#'s are in one col
        tmp <- pivot_longer(data = tmp,
                            cols = which(startsWith(colnames(tmp), "Mouse_")),
                            names_to = "Mouse_num",
                            values_to = "tumor_vol")
        #Remove "Invalid Number", set cutoff of 28 days and calculate relative_tumor_vol # nolint
        tmp <- tmp %>%
            filter(tumor_vol != "Invalid Number" &
                   timepoint <= 28) %>%
            group_by(Mouse_num) %>%
            mutate(relative_tumor_vol = tumor_vol/tumor_vol[1])
        #Add this sheet to dataframe with info for this experiment
        sheet_df <- rbind(sheet_df, tmp)
    }
    #Add experiment data frame to list of dataframes
    dfs_list[[i]]  <- sheet_df
}
```

## Generating Plots

Now that the data is loaded in and organized, I will generate a waterfall plot displaying best overall response for each experiment.
```{r}
#Generate Waterfall plots for all experiments
summarized_dfs_list <- lapply(dfs_list, calc_resp)
summarized_dfs_list <- lapply(summarized_dfs_list, summarized_df)
plots <- lapply(summarized_dfs_list, make_wtrfall_plot)
pdf("Figures/waterfall_plots.pdf")
print(plots)
dev.off()
```

## Pairwise Fisher Tests

For all experiments we will test the following null hypothesis at a significance level of 0.05 between each pair of treatments: the objective response rate is the same between the two treatments.
```{r}
#Create table of responses and then perform pairwise fisher multiple tests
fisher_tests <- lapply(summarized_dfs_list, function(x) {
    tmp <- group_by(x, Treatment) %>%
        summarize("Response" = sum(calc_resp %in% c("CR", "PR", "SD")),
                  "No_Response" = sum(calc_resp == "PD")) %>%
        column_to_rownames(var = "Treatment")
    rstatix::pairwise_fisher_test(tmp, p.adjust.method = "BH")
})
write_xlsx(
    x = fisher_tests,
    path = "Output/fisher_test.xlsx"
)

#Generate list of tables with responses for each treatment in each experiment
resp_tabs <- lapply(summarized_dfs_list, function(x) {
    group_by(x, Treatment) %>%
        summarize("Response" = sum(calc_resp %in% c("CR", "PR", "SD")),
                  "No_Response" = sum(calc_resp == "PD")) %>%
        column_to_rownames(var = "Treatment")
})
```